---
title: "Dispersed urban-stormwater control improved stream water quality in a catchment-scale experiment"
subtitle: "Supplementary methods E: model compilation with predictions to new data"
author:
  - Christopher J Walsh, Sam J Imberger, Matthew J Burns, Darren G Bos, and Tim D Fletcher
date: "School of Ecosystem and Forest Sciences, The University of Melbourne, 500 Yarra Boulevard, Burnley, 3121 Victoria, Australia"
output:   
  word_document:
    reference_docx: officedown_template.docx
csl: ecol-app.csl
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, 
                      message = FALSE, error = FALSE)
library(rstan); library(loo)
rstan_options(auto_write = TRUE)
# Set a directory for storing the (large) model objects: 
# I used a directory on a network drive
# (the directory needs three sub-directories: data, model_fits and compiled_models)
mod_dir <- "[Network Drive]/git-data/lsc_dbs_wq/" #
```

This document compiles and saves the Stan model objects described in `wq_manuscript.Rmd`, sampled in
`wq_model_sampling.Rmd` to model the water quality data compiled in `wq_data_compilation.Rmd`. The models are the same as those compiled in `wq_model_compilation.Rmd`, except they include prediction to a new dataset in the generated quantities section, for production of counterfactual plots of model predictions.

```{stan output.var="m_1_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p [N2];
    real restr_p [N2];
    real rain1_p [N2];
    real t_p [N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    vector[N_site_no] b_t;
    real t_mu;
    real<lower=0> t_sigma;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    t_mu ~ normal( 0, 5 );
    t_sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( t_mu , t_sigma );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + t_mu * t_p[i]; 
    }
}

```

```{r}
save(m_1_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1_st_p.rda"))
```

```{stan output.var="m_1_x_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real add1[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p [N2];
    real restr_p [N2];
    real rain1_p [N2];
    real add1_p [N2];
    real t_p [N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_add1;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    vector[N_site_no] b_t;
    real t_mu;
    real<lower=0> t_sigma;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    t_mu ~ normal( 0 , 5 );
    t_sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( t_mu , t_sigma );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_add1 ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                b_add1 * add1_p[i] + 
                b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + t_mu * t_p[i]; 
    }
}
```

```{r}
save(m_1_x_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1_x_st_p.rda"))
```

```{stan output.var="m_1_xx_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real add1[N];
    real add2[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p [N2];
    real restr_p [N2];
    real rain1_p [N2];
    real add1_p [N2];
    real add2_p [N2];
    real t_p [N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_add1;
    real b_add2;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    vector[N_site_no] b_t;
    real t_mu;
    real<lower=0> t_sigma;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    t_mu ~ normal( 0 , 5 );
    t_sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( t_mu , t_sigma );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_add1 ~ normal( 0 , 5 );
    b_add2 ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t[site_no[i]] * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                b_add1 * add1_p[i] + b_add2 * add2_p[i] + 
                b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + t_mu * t_p[i]; 
    }
}
```

```{r}
save(m_1_xx_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1_xx_st_p.rda"))
```

```{stan output.var="m_1ft_xx_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real add1[N];
    real add2[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p [N2];
    real restr_p [N2];
    real rain1_p [N2];
    real add1_p [N2];
    real add2_p [N2];
    real t_p [N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_add1;
    real b_add2;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    real b_t;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( 0, 5 );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_add1 ~ normal( 0 , 5 );
    b_add2 ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                b_add1 * add1_p[i] + b_add2 * add2_p[i] + 
                b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + b_t * t_p[i]; 
    }
}
```

```{r}
save(m_1ft_xx_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1ft_xx_st_p.rda"))
```

```{stan output.var="m_1ft_xxx_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real add1[N];
    real add2[N];
    real add3[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p [N2];
    real restr_p [N2];
    real rain1_p [N2];
    real add1_p [N2];
    real add2_p [N2];
    real add3_p [N2];
    real t_p [N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_add1;
    real b_add2;
    real b_add3;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    real b_t;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( 0 , 5 );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_add1 ~ normal( 0 , 5 );
    b_add2 ~ normal( 0 , 5 );
    b_add3 ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                b_add1 * add1_p[i] + b_add2 * add2_p[i] + b_add3 * add3_p[i] + 
                b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + b_t * t_p[i]; 
    }
}
```

```{r}
save(m_1ft_xxx_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1ft_xxx_st_p.rda"))
```


```{stan output.var="m_3ft_sdr_int_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real season[N];
    real diel[N];
    real rain365[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p[N2];
    real restr_p[N2];
    real rain1_p[N2];
    real season_p[N2];
    real diel_p[N2];
    real rain365_p[N2];
    real t_p[N2];
}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_seas;
    real b_diel;
    real b_seas_r;
    real b_seas_d;
    real b_diel_r;
    real b_diel_d;
    real b_auto_s;
    real b_auto_t;
    real b_rain365;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    real b_t;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( 0 , 5 );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_seas ~ normal( 0 , 5 );
    b_diel ~ normal( 0 , 5 );
    b_rain365 ~ normal( 0, 5 );
    b_seas_r ~ normal( 0 , 5 );
    b_seas_d ~ normal( 0 , 5 );
    b_diel_r ~ normal( 0 , 5 );
    b_diel_d ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i]  + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + b_rain365 * rain365[i] + 
                b_seas * season[i] + b_diel * diel[i] + b_seas_r * season[i] * restr[i] + 
                b_seas_d * season[i] * degrd[i] +  b_diel_r * diel[i] * restr[i] + 
                b_diel_d * diel[i] * degrd[i] + b_p * rain1[i] + b_dp * degrd[i] * rain1[i] +
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }

    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );

    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + b_rain365 * rain365_p[i] + 
        b_seas * season_p[i] + b_diel * diel_p[i] +  b_seas_r * season_p[i] * restr_p[i] + 
                b_seas_d * season_p[i] * degrd_p[i] +  b_diel_r * diel_p[i] * restr_p[i] + 
                b_diel_d * diel_p[i] * degrd_p[i] + b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i]  + 
                b_rp * restr_p[i] * rain1_p[i] + b_t * t_p[i]; 
    }

}
```

```{r eval = TRUE}
save(m_3ft_sdr_int_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_3ft_sdr_int_st_p.rda"))
```


```{stan output.var="m_1ft_xxx_int_st_p"}
data{
    int<lower=1> N;
    int<lower=1> N_site_no;
    real y[N];
    int site_no[N];
    real degrd[N];
    real restr[N];
    real rain1[N];
    real add1[N];
    real add2[N];
    real add3[N];
    real t[N];
    real auto_t[N]; // the temporal autocorrelation term
    real auto_s[N]; // the spatial autocorrelation term
    int n_st_miss;  // the number of data points missing both auto_s and auto_t
    int n_s_miss;   // the number of data points missing auto_s (but not auto_t)
    int n_t_miss;   // the number of data points missing auto_t (but not auto_s)
    int n_no_miss;  // the number of data points with both auto_t and auto_s values
    int st_miss[n_st_miss]; // indices of data points missing both auto_s and auto_t
    int s_miss[n_s_miss];   // indices of data points missing auto_s (but not auto_t)
    int t_miss[n_t_miss];   // indices of data points missing auto_t (but not auto_s)
    int no_miss[n_no_miss]; // indices of data points with both auto_t and auto_s values
    int<lower=1> N2;
    real degrd_p[N2];
    real restr_p[N2];
    real rain1_p[N2];
    real add1_p[N2];
    real add2_p[N2];
    real add3_p[N2];
    real t_p[N2];

}
parameters{
    real a;
    real b_d;
    real b_r;
    real b_p;
    real b_dp;
    real b_rp;
    real b_add1;
    real b_add1p;
    real b_add2;
    real b_add3;
    real b_auto_s;
    real b_auto_t;
    vector[N_site_no] a_s;
    real<lower=0> a_s_sigma;
    real b_t;
    real<lower=0> sigma;
    }
model{
    vector[N] mu;
    sigma ~ cauchy( 0 , 2 );
    b_t ~ normal( 0 , 5 );
    a_s_sigma ~ cauchy( 0 , 2 );
    a_s ~ normal( 0 , a_s_sigma );
    b_r ~ normal( 0 , 5 );
    b_d ~ normal( 0 , 5 );
    b_p ~ normal( 0 , 5 );
    b_dp ~ normal( 0 , 5 );
    b_rp ~ normal( 0 , 5 );
    b_add1 ~ normal( 0 , 5 );
    b_add1p ~ normal( 0 , 5 );
    b_add2 ~ normal( 0 , 5 );
    b_add3 ~ normal( 0 , 5 );
    b_auto_s ~ normal( 0 , 5 );
    b_auto_t ~ normal( 0 , 5 );
    a ~ normal( 0 , 10 );
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    y ~ normal( mu , sigma );
}
generated quantities{
    vector[N] mu;
    vector[N] log_lik;
    vector[N2] y_pred;
    for ( i in st_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i];
    }
    for ( i in s_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] +  
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i];
    }
    for ( i in t_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] + b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i]  + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_s * auto_s[i];
    }
    for ( i in no_miss ) {
        mu[i] = a + a_s[site_no[i]] + b_d * degrd[i] + b_r * restr[i] + 
                b_add1 * add1[i] + b_add2 * add2[i] +  b_add3 * add3[i] + 
                b_p * rain1[i] + b_dp * degrd[i] * rain1[i] + b_add1p * add1[i] * rain1[i] + 
                b_rp * restr[i] * rain1[i] + b_t * t[i] + 
                b_auto_t * auto_t[i]  + b_auto_s * auto_s[i];
    }
    for ( i in 1:N ) log_lik[i] = normal_lpdf( y[i] | mu[i] , sigma );
    for ( i in 1:N2 ) {
        y_pred[i] = a + b_d * degrd_p[i] + b_r * restr_p[i] + 
                    b_add1 * add1_p[i] + b_add2 * add2_p[i] +  b_add3 * add3_p[i] + 
                    b_p * rain1_p[i] + b_dp * degrd_p[i] * rain1_p[i] + b_add1p * add1_p[i] * rain1_p[i] +                
                    b_rp * restr_p[i] * rain1_p[i] + b_t * t_p[i]; 
     }
    }

```

```{r}
save(m_1ft_xxx_int_st_p, file = paste0(mod_dir, "compiled_models/wq_stan_m_1ft_xxx_int_st_p.rda"))
```



